---
title: "Capstone Data Visualization"
author: "Zahra Nur Anisah"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
output:
  html_document:
    theme: cosmo
    highlight: breezedark
    toc: true
    toc_float:
      collapsed: false
    df_print: paged
---

```{r setup, include=FALSE}
# clear-up the environment
rm(list = ls())

# chunk options
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  comment = "#>"
)

options(scipen = 100)
```


Goal: Capstone Dashboard

Tahapan: 
- Preprocessing 
  - Cleansing
  - Perubahan tipe data
  - Feature Engineering
  - Aggregasi data
- Buat Tooltip
- Visualisasi Data
- Buat visualisasi interaktif
- Web dashboard


  - Tahapan preprocessing/ wrangling:
    - ✅Subset kolom ~> select()
    - ✅Filter baris  ~>  filter()
    - ✅Mengubah tipe data ~> mutate()
    - ✅Menambah kolom baru ~> mutate()
    - ✅Aggregasi data ~>  group_by() + summarise() + ungroup()
    - ✅Mengurutkan data ~>  arrange()
  
  
# Introduction

```{r, warning=FALSE}
library(dplyr)
library(ggplot2) # visualisasi
library(ggpubr)
library(scales) # untuk tampilan digit (memberikan koma dll)
library(glue)
library(plotly) 
library(lubridate) # working with datetime
options(scipen = 100) # supaya output tidak menampilkan notasi ilmiah (10-e10)
```

# Load Data

Kita akan coba mengeksplorasi kembali data **Pakistan Largest Ecommerce Dataset** menggunakan package dplyr! Mari kita baca terlebih dahulu data kita:

```{r}
# read data
data <- read.csv("Pakistan Largest Ecommerce Dataset.csv", stringsAsFactors = TRUE, encoding = "latin1")

# cleansing data
data_clean <- data %>% 
  select(
    #membersihkan yang tidak perlu
    -c(increment_id, sales_commission_code, Working.Date, BI.Status,MV, Year, Month, Customer.Since, M.Y, FY, X, X.1, X.2, X.3, X.4)) %>% 
  
  mutate(
    #ngubah tipe data
    created_at= dmy(created_at),
    sku= as.character(sku),
    
    #nambah kolom
    month= month(created_at, label = FALSE, abbr = TRUE),
    year= year(created_at)
  ) %>% 
  
  #gantii nama kolom
  rename(
    c(product_name = sku, category_name = category_name_1)) %>% 

  filter(
    #filter hanya menggunakan tahun 2017
    year(created_at) %in% c(2017),
    category_name != "\\N",
    grepl("_", product_name) # filter nama produk yang tidak adaa "_" atau kelihatan seperti barcode 
  )

data_clean
```

```{r}
# cek apabila ada missing data
data_clean %>% is.na() %>% colSums()
```
Poin captone:
- ketepatan kita menggunakan plot dalam membandingkan data yang ingin kitaa representasikan 

#Page 1(Overview)
##PLOT 1: (line plot)
"Monthly Sales Trend in Pakistan for 2017"
```{r}
# Data Wrangling
sales_trend <- data_clean %>% 
  group_by(month) %>% 
  summarise(total_sales = sum(grand_total)) %>% 
  ungroup() %>% 
  arrange(month)
sales_trend
```
```{r}
# nambah kolom month singkatan & full month ~> utk x axis di plot & tooltip
sales_trend <- sales_trend %>%
  mutate(
    # kolom singkatan utk sumbu x
    month_name = factor(month, levels = 1:12, labels = month.abb),
    # kolom full month utk tooltip
    full_month_name = factor(month, levels = 1:12, labels = month.name)
  )
sales_trend
```

```{r}
#buat tooltip desc
sales_trend <- sales_trend %>% 
  mutate(
    label = glue(
      "Total Sales: {comma(total_sales)}
      Month: {full_month_name}"
    )
  )
sales_trend
```


```{r}
# Visualization ~> dibuat ggplot nyaa
#"Monthly Sales Trend in Pakistan for 2017"

plot1 <- ggplot(sales_trend, aes(x=month_name, y= total_sales))+
  
  geom_line(col="darkgreen", group=1) +
  geom_point(aes(text=label), col="black") +
  scale_y_continuous(labels = comma, breaks = seq(0, 100000000, 10000000)) +
  labs(
    title = "Monthly Sales Trend",
    x = NULL,
    y = "Total Sales"
  ) +
  theme_minimal() +
  theme(
     plot.title = element_text(hjust = 0.5)
  )

ggplotly(plot1, tooltip = "text")

```
> This shows the eCommerce sales growth throughout the year.

## Total Sales
```{r}
data_clean
```

```{r}
Total_Sales<- data_clean %>% summarise(Total_sales=comma(sum(grand_total)))

Total_Sales
```

## Total Orders
```{r}
Total_Orders<- data_clean %>% summarise(Total_orders=n())

Total_Orders
```
## Total Products
```{r}
Total_product<- data_clean %>% summarise(total_product=comma(sum(qty_ordered)))

Total_product
```
## Total Customer
```{r}
Total_customer<- data_clean %>% 
  summarise(total_cust=comma(length(unique(Customer.ID))))

Total_customer
```


##PLOT 2: (lolipop plot)
"Total Orders by Payment Method in Pakistan"

```{r}
# Data Wrangling 
fav_payment<- data_clean %>% 
  group_by(payment_method) %>% 
  summarise(Payment_type = n()) %>% 
  ungroup() %>% 
  arrange(-Payment_type) 

fav_payment
```


```{r}
fav_payment <- fav_payment %>% 
  mutate(
    label = glue(
      "Payment Method: {payment_method}
      Total Orders: {comma(Payment_type)}"
    )
  )

fav_payment
```


```{r}
# Cara 1: INI PAKE LOLIPOP PLOT, efektif

# Visualization ~> dibuat ggplot nyaa
#"Fav Payment Method"
plot2 <- ggplot(fav_payment, aes(x = Payment_type, 
                    y = reorder(payment_method, Payment_type),
                    text = label)) +
  
  geom_segment(aes(x= 0, xend=Payment_type, yend=reorder(payment_method, Payment_type)), color="darkgreen", size= 1) +
  geom_point(color="black", size=3) +
  scale_x_continuous(labels = comma) +
  labs(title = "Top Payment Method",
       x = "Total Orders",
       y = NULL) +
  theme_minimal() +
  theme(
     plot.title = element_text(hjust = 0.5)
  )

ggplotly(plot2, tooltip = "text")
```

```{r}
# Cara 2: INI PAKE BAR PLOT, tp susah utk buka tooltip rank paling rendah krn qty dikit jd barnya jg tipis

# Visualization ~> dibuat ggplot nyaa
#"Fav Payment Method"
plot2 <- ggplot(fav_payment, aes(x = Payment_type, 
                    y = reorder(payment_method, Payment_type),
                    text = label)) +
  
  geom_col(aes(fill=Payment_type)) +
  scale_x_continuous(labels = comma)+
  scale_fill_gradient(low="black", high="darkgreen") +
  labs(title = "Top Payment Method in Pakistan 2017",
       x = "Total Orders",
       y = NULL) +
  theme_minimal()+
  theme(legend.position = "none")

ggplotly(plot2, tooltip = "text")
```
>This identifies the most popular payment methods in the country.


#Page 2(Sales Analysis)
## PLOT 3: (line plot)
"Monthly total sales in each category"
```{r}
data_clean
```

```{r}
# Data Wrangling
category_sales_trend <- data_clean %>%
  filter(category_name =="Beauty & Grooming") %>% 
  group_by(month) %>% 
  summarise(total_sales_category = sum(grand_total)) %>% 
  ungroup() %>% 
  arrange(month)

category_sales_trend
```

```{r}
# nambah kolom month singkatan & full month ~> utk x axis di plot & tooltip
category_sales_trend <- category_sales_trend %>%
  mutate(
    # kolom singkatan utk sumbu x
    month_name = factor(month, levels = 1:12, labels = month.abb),
    # kolom full month utk tooltip
    full_month_name = factor(month, levels = 1:12, labels = month.name)
  )

category_sales_trend
```

```{r}
#buat tooltip desc
category_sales_trend <- category_sales_trend %>% 
  mutate(
    label = glue(
      "Total Sales: {comma(total_sales_category)}
      Month: {full_month_name}"
    )
  )

category_sales_trend
```


```{r}
# CARA 1: INI DGN LINE PLOT, GAMPANG DIBACA

# Visualization ~> dibuat ggplot nyaa
#"Monthly total sales in each category"

plot3 <- ggplot(category_sales_trend, aes(x=month_name, y= total_sales_category))+
  
  geom_line(col="darkgreen", group=1) +
  geom_point(aes(text=label), col="black") +
  scale_y_continuous(labels = comma, breaks = seq(0, 100000000, 200000)) +
  labs(
    title = "Monthly Sales for Beauty & Grooming",
    x = NULL,
    y = "Total Sales"
  ) +
  theme_minimal() +
  theme(
     plot.title = element_text(hjust = 0.5)
  )

ggplotly(plot3, tooltip = "text")
```

```{r}
# CARA 2: INI DGN LOLIPOP PLOT

# Visualization ~> dibuat ggplot nyaa
#"Monthly total sales in each category"
plot3 <- ggplot(category_sales_trend, aes(x=month_name, y= total_sales_category, text=label)) +
  
  geom_segment(aes(xend=month_name, y=0,yend=total_sales_category), color="darkgreen", size= 1) +
  geom_point(color="black", size=3) +
  scale_y_continuous(labels = comma, , breaks = seq(0, 100000000, 200000)) +
  labs(title = "Monthly Total Sales on Beauty & Grooming in Pakistan 2017",
       x = NULL,
       y = "Total Sales") +
  theme_minimal() +
  theme(
     plot.title = element_text(hjust = 0.5)
  )

ggplotly(plot3, tooltip = "text")
```
>This gives a view of which month was has the most sales in Pakistan.


##PLOT 4: (bar plot)
"Top 10 Best Selling Product in each category"
```{r}
data_clean
# unique(data_clean$payment_method) #cek jenis pembayaran
```

```{r}
# Data Wrangling
fav_product <- data_clean %>% 
  filter(category_name =="Mobiles & Tablets") %>% 
  group_by(product_name) %>% 
  summarise(total_qty = sum(qty_ordered), total_sold= sum(grand_total)) %>% 
  ungroup() %>% 
  arrange(-total_qty) %>% 
  head(10)

fav_product
```
```{r}
process_product_name <- function(product_name) {
  parts <- str_split(product_name, " - ") %>% unlist()  # Split by ' - '
  if (length(parts) > 2) {
    # Combine the first and last part (brand name & color)
    return(paste(parts[1], parts[length(parts)], sep = " - "))
  } else {
    # If there are not 3 parts, return the product name as is
    return(product_name)
  }
}
```


```{r}
library(stringr)
fav_product<- fav_product %>% 
  mutate(short_name = sapply(product_name, process_product_name)) %>% # Truncate to 20 chars
  mutate(label = glue(
      "Product : {product_name}
      Total Sales : {comma(total_sold)}
      Quantity : {total_qty}"
    )
 )

fav_product
```

```{r}
# Visualization ~> dibuat ggplot nyaa
# "Top 10 Best Selling Product in each category"
plot4 <- ggplot(fav_product, aes(x = total_sold,
                    y = reorder(short_name, total_sold),
                    text = label)) +
  
  geom_col(aes(fill = total_sold)) +
  scale_fill_gradient(low="black", high="darkgreen") +
  labs(title = "Top Mobiles & Tablets Products",
       x = "Total Sales",
       y = NULL) +
  scale_x_continuous(labels = comma) +
  theme_minimal() +
  theme(legend.position = "none",  plot.title = element_text(hjust = 0.5))

ggplotly(plot4, tooltip = "text")
```
> This shows the most popular product by the number of items sold & contributed to the highest sales in Pakistan.


# Page 3(Customer Preferance)
##PLOT 5: (bar plot)
"Quantity Product in each category Ordered over the year 2017"
```{r}
data_clean
```

```{r}
# Data Wrangling
rank_qty_pcategory <- data_clean %>% 
  filter(created_at >= as.Date("2017-01-01") & created_at <= as.Date("2017-01-06")) %>% 
  group_by(category_name) %>% 
  summarise(qty_pcategory = sum(qty_ordered)) %>% 
  ungroup() %>% 
  arrange(-qty_pcategory) 

rank_qty_pcategory
```


```{r}
#buat desc tooltip
rank_qty_pcategory <- rank_qty_pcategory %>% 
  mutate(label = glue(
      "Category : {category_name}
      Quantity : {comma(qty_pcategory)}"
    )
 )

rank_qty_pcategory
```

```{r}
# Visualization ~> dibuat ggplot nyaa
# Quantity Product in each category Ordered over the year 2017"
plot5 <- ggplot(rank_qty_pcategory, aes(x = qty_pcategory,
                    y = reorder(
category_name, qty_pcategory),
                    text = label)) +
  
  geom_col(aes(fill = qty_pcategory)) +
  scale_fill_gradient(low="black", high="darkgreen") +
  labs(title = "Top Selling Categories",
       x = "Quantity",
       y = NULL) +
  scale_x_continuous(labels = comma) +
  theme_minimal() +
  theme(legend.position = "none",  plot.title = element_text(hjust = 0.5))

ggplotly(plot5, tooltip = "text")
```
>  This shows the most popular categories based on the qty from a certain range of date.


##PLOT 6: (scatter plot)
"Relationship between price and quantity from a certain range of date"
```{r}
data_clean
```


```{r}
# Data Wrangling
rank_qty_pprice <- data_clean %>% 
  filter(created_at >= as.Date("2017-01-01") & created_at <= as.Date("2017-01-06")) %>% 
  group_by(price) %>% 
  summarise(qty_pprice = sum(qty_ordered)) %>% 
  ungroup() %>% 
  arrange(-qty_pprice) 

rank_qty_pprice
```


```{r}
#buat desc tooltip
rank_qty_pprice <- rank_qty_pprice %>%
  mutate(label = glue(
      "Price : {comma(price)}
      Quantity : {comma(qty_pprice)}"
    )
 )

rank_qty_pprice
```


```{r}
# Visualization ~> dibuat ggplot nyaa
# Quantity Product in each category Ordered over the year 2017"
plot6 <- ggplot(rank_qty_pprice, aes(x = qty_pprice,
                    y = price,
                    text = label)) +
  
  geom_jitter(color="darkgreen") +
  labs(title = "Relationship Between Price and Quantity Sold Products",
       x = "Quantity",
       y = "Price") +
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = comma)+
  theme_minimal() +
  theme(
     plot.title = element_text(hjust = 0.5)
  )

ggplotly(plot6, tooltip = "text")
```
> Shows the relationship between price and quantity. The higher the price, the less it is sold, and the higher it was sold, the lower the price are.


##PLOT 7: (line plot)
"Sales/ Trend over date with many categories(top 8)"
```{r}
data_clean
```

```{r}
# Data Wrangling
rank_sales_cat <- data_clean %>% 
  filter(created_at >= as.Date("2017-01-01") & created_at <= as.Date("2017-01-06")) %>% 
  group_by(category_name) %>% 
  summarise(sales = sum(grand_total)) %>% 
  ungroup() %>% 
  arrange(-sales) 

rank_sales_cat
```

```{r}
rank_sales_cat <- rank_sales_cat %>% 
  mutate(
    label = glue(
      "Category: {category_name}
      Total Sales: {comma(sales)}"
    )
  )

rank_sales_cat
```

```{r}
# Visualization ~> dibuat ggplot nyaa
#"Top Categories by Sales"
plot7 <- ggplot(rank_sales_cat, aes(x = sales, 
                    y = reorder(category_name, sales),
                    text = label)) +
  
  geom_segment(aes(x= 0, xend=sales, yend=reorder(category_name, sales)), color="darkgreen", size= 1) +
  geom_point(color="black", size=3) +
  scale_x_continuous(labels = comma) +
  labs(title = "Top Category Sales",
       x = "Total Sales",
       y = NULL) +
  theme_minimal() +
  theme(
     plot.title = element_text(hjust = 0.5)
  )

ggplotly(plot7, tooltip = "text")
```
>  This ranked the highest sales categories from a certain range of date.

#DONE